sensi_an_rotations1. Start of sensitivity analysis for rotations.

### 2018-05-18 inst/documents/paper3/pap3_rot_analyses.Rmd
paper3

  # load simulation outputs for some plots
  # beware that each object is called df_res_all
  # some plots are created by directly running the model
  load("df_res_all_rot_nocost_nomig1000.rda") #nocost nomig  
  df_nocost_nomig <- df_res_all
  load("df_res_all_rot10000.rda") #cost
  df_cost <- df_res_all
  load("df_res_all_rot_mig10000.rda") #mig 10k
  df_mig <- df_res_all
  load("df_res_all_rot_cost_mig10000.rda") #costmig 10k
  df_cost_mig <- df_res_all

@@

>>>>

  # load simulation outputs for some plots
  # beware that each object is called df_res_all
  # some plots are created by directly running the model
  load("df_res_all_rot_nocost_nomig1000.rda") #nocost nomig  
  df_nocost_nomig <- df_res_all
  load("df_res_all_rot10000.rda") #cost
  df_cost <- df_res_all
  load("df_res_all_rot_mig10000.rda") #mig 10k
  df_mig <- df_res_all
  load("df_res_all_rot_cost_mig10000.rda") #costmig 10k
  df_cost_mig <- df_res_all
  load("df_res_all_rot_nocost_nomig10000_mgs5.rda")
  df_nocost_nomig_mgs5 <- df_res_all 
  load("df_res_all_rot_cost10000_mgs5.rda")
  df_cost_mgs5 <- df_res_all

@@

\pagebreak
Showing that using min_gens_switch_back=5 instead of min_rwr_interval=5 seems to make no difference in main results.

```{r min_gens_switch_back on success, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9}

# if I put all scenarios in one df with a column for scenario I can do one ggplot and facet 
# would work for the previous plot too and would simplify adding the new columns
# first add the scenario column
df_nocost_nomig$scenario <- 'no cost or dispersal'
df_cost$scenario <- 'costs'
df_mig$scenario <- 'dispersal'
df_cost_mig$scenario <- 'costs and dispersal'

df_nocost_nomig_mgs5$scenario <- 'no cost or dispersal mgs5'
df_cost_mgs5$scenario <- 'costs mgs5'

df_all <- dplyr::bind_rows(df_nocost_nomig,
                           df_nocost_nomig_mgs5,
                           df_cost,
                           df_mig,
                           df_cost_mig,
                           df_cost_mgs5)
#order for ggplot
df_all$scenario <- factor(df_all$scenario, levels=c('no cost or dispersal','no cost or dispersal mgs5','costs','costs mgs5','dispersal','costs and dispersal'))

# add a column for classification of whether, both, rotation, or neither succeeded
df_all <- mutate(df_all, success = ifelse(gens_seq==max_gen & gens_rot==max_gen,'both',
                                               ifelse(gens_rot==max_gen,'rotation only',
                                               ifelse(gens_seq==max_gen,'sequence only', 'neither'  
                                                      ))))
#count(df_all, scenario, success)

#hack to add subplot labels to facets
abcd <- data.frame(label = c("A","B","C","D","E","F"),
                   scenario = levels(df_all$scenario))

ggplot( df_all, aes(x=success)) + 
        #geom_bar() +
        stat_count(mapping = aes(y=..prop.., group=1)) + #puts proprtions in axes
        scale_y_continuous(breaks=seq(0.2,0.8,0.2)) +
        theme_minimal() +
        ylab('proportion of model runs') +
        facet_wrap(~scenario, ncol=1) + #, scales='free_y') + #can remove free_y later if I do 10k nocost runs    
        geom_text(data=abcd,
                  aes(x = -Inf, y = +Inf, label=label),
                  hjust = 0,
                  vjust = 0.9 )

\pagebreak
Using min_gens_switch_back=5 gives greater benefits to sequences when there are no costs or dispersal. I suspect this is an artefact. Try looking at some of the runs.
```{r min_gens_switch_back on rotation benefit, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9}

  #try quicker way with df_all and facet
  df_all <- mutate(df_all, pcent_rot_adv = rot_minus_seq * 100/gens_seq)
  custom_bins <- c(-20,-10,-1,1,10,20,100,450)
  df_all$rot_adv_bins <- cut(df_all$pcent_rot_adv, custom_bins)  
  bin_labs <- levels(df_all$rot_adv_bins)

  #hack to add subplot labels to facets
  abcd <- data.frame(label = c("A","B","C","D","E","F"),
                   scenario = levels(df_all$scenario))

  ggplot( df_all, aes(x=rot_adv_bins)) +

             #geom_bar() +
             stat_count(mapping = aes(y=..prop.., group=1)) + #puts proprtions in axes
             scale_y_continuous(breaks=seq(0,0.8,0.2)) +
             xlab('Percent benefit of rotation over sequence') +
             ylab('proportion of model runs') +
             #ggtitle('') +
             theme_minimal() +
             facet_wrap(~scenario, ncol=1) +
             scale_x_discrete( limits=bin_labs,
                               labels=c("-20 to -10", "-10 to -1", "-1 to 1", "1 to 10", "10 to 20", "20 to 100", "100 to 500"))+     
             theme(panel.grid.minor.y=element_blank(),
                   axis.text.x = element_text(angle = 20,hjust = 1, vjust = 1)) +
             geom_text(data=abcd,
                  aes(x = -Inf, y = +Inf, label=label),
                  hjust = 0,
                  vjust = 0.9 )

```
\pagebreak
Using min_gens_switch_back=5 example of when sequence does better.
```{r min_gens_switch_back sequence better example, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9}

df_seqbetter <- filter( df_all, pcent_rot_adv < -5 &
                                scenario == 'no cost or dispersal mgs5' )

#ooo scenario 5 has rot_adv of -97%
#but remember it is not scenario5 in the original dataset !
#Aha! all these big sequence advantages (well those >50%) 
#were due to artfefact when rotation interval=5 and num insecticides=2 rotation couldn't return to first insecticide because it was only 5 generations ago !!

df_nocost_nomig_mgs5 <- filter( df_all, scenario == 'no cost or dispersal mgs5' )

#which(df_nocost_nomig_mgs5$pcent_rot_adv < -90)
#217 1933 3435 4289 5649 7100 7925 7990

scenarios_to_plot <- c(217)

for(scenario in scenarios_to_plot)
{
  #cat(scenario,"\n")

  linputs <- as.list( df_nocost_nomig_mgs5[scenario, 1:15] )

  # run one scenario
  dfres <- do.call(run_rot, linputs)

  ggrot <- rot_plot_resistance(dfres, plot_refuge=FALSE, logy=TRUE, plot=FALSE)

  #now run as sequence
  linputs$rot_interval <- 0
  dfres <- do.call(run_rot, linputs)

  ggseq <- rot_plot_resistance(dfres, plot_refuge=FALSE, logy=TRUE, plot=FALSE)

  grid.arrange(ggrot, ggseq, ncol=1)

}

### 2018-05-18 inst/documents/paper3/paper3_rotations.Rmd
paper3

*todo add to this description of how the model works and refer to SI where we can show the algebra*    
Within the model, the inputs described in Table 1 determine the fitness in each generation as shown in Figure 1, differences in fitness lead to the changes in allele frequencies in the following generation.

@@

*Ian : at this stage it would be good to revisit these model details before we continue further.*    
*It is fairly easy for me to rerun all the analyses if we change our mind on these.*

#### Optional model details    
Two optional model details were included to aid giving a more meaningful comparison of strategies. Firstly allele frequencies in a run were not allowed to go below the starting frequencies randomly chosen for that run. This aims to represent a situation where in the absence of insecticide use allele frequencies are maintained in balance between mutation and selection. Without this option enabled it was found that

Secoundly a constraint was added to stop rapid switching between insecticides every few generations which would be impractical in the field. Certain input combinations create this behaviour in the absence of a constraint. For example when two insecticides are close to the resistance frequency threshold of 0.5 it is possible for them to switch between each other every few generations as frequencies go above the threshold when in use and decline below it when not in use.  
*todo finish this part*

@@

*todo change this bit, think more about which outputs to use*
It is necessary to choose a measure to quantify the performance of an insecticide-use strategy and enable comparison with an alternative. For this we summed the number of generations when an insecticide was deployed and the resistance allele frequency for that insecticide was below 0.5. This allowed comparison between strategies using a variable number of insecticides. Previously time-to-resistance, namely the number of generations it takes to reach a defined resistance threshold (usually a resistance allele frequency of 0.5), was used. Using time-to-resistance worked when considering just two insecticides, but when the number of insecticides is a variable it is not sufficient.

@@

*todo continue describing the example runs first, before talking about overall results*

### Results2 : comparing rotations and sequences across model runs.

@@

Thus to aid the comparison of rotations and sequences model runs can be broadly classified into four groups :
1. both strategies succeeded : for both rotations and sequences resistance thresholds for all insecticides not reached within 500 generations.
2. neither strategy succeeded
3. rotation only succeeded
4. sequence only succeeded

The number of model runs in each of these groups for the different scenarios are shown in Fig 5. In the base scenario with no costs or dispersal (Fig 5A) the majority of runs do not succeed, a small proportion succeed for both strategies and none succeed for either strategy alone. When costs of reistance are added (Fig 5B) more runs succeed for both strategies, up to about half of those that succeed for neither, and a small proportion (< 4%) succeed for rotations alone. Adding both costs and dispersal results in more runs succeeding for both than neither (Fig 5C).

@@

*refer to violin plots here ?*

As shown in Fig 4.2 some runs with costs of resistance can show substantial advantages for rotations over sequences. Examples of one such run is shown in Fig 5.

A Partial Rank Correlation analysis identified that of the inputs, dominance of cost and dominance of selection had the greatest influence on the difference between rotations and sequences when resistance costs were included (Fig 6). A combination of high values of dominance of cost and low values for dominance of selection are required to give the greatest advantages to rotations, but do not guarantee this advantage (Fig 7).

@@

This can be understood intuitively. When there are no costs, the dynamics are the same, it is simply that sequential deployment proceeds in a few, larger steps while rotations proceeds in more numerous smaller steps; both reach the same point in around the same time. However dominance effects on cost and resistance can act to keep frequencies low in a rotations policy. Recall that at low frequencies of resistance most resistance alleles are present as heterozygotes RS and only a vey small proportions heterozygotes. However costs are dominant and still act at low frequencies. This is illustrated on Figure Y. Relatively rapid rotations keep the allele frequencies at the LH side of Figure Y where costs during the periods when the insecticide is not deployed greatly outweigh the selection for IR during the periods when the insecticide is deployed. We believe the explanation is therefore that rotations keep allele frequencies in the parameter space to the left of Figure Y and hence acts to slow the evolution of IR compared to sequences which allow frequencies to rise to relatively high levels before the insecticide is replaced.  (does this sound reasonable??). Note also that starting frequency, as expected, had an impact (Figure 7), lower starting frequencies slightly favouring rotations by allowing more time in the “low frequency” parameter space.

@@

1. *add to methods description of how fitness determined in model, move & modify fig 1*
1. *add caveat about polygenic resistance*

@@

@@

```{r Fig6, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=4, fig.show='hold', fig.align='center', fig.cap="Figure 6. Effect of cost and dispersal on the difference between rotations and sequences from model runs. The y axis shows the number of model runs for the different caegories of difference between rotations and sequences. A) when there are no resistance costs or dispersal from untreated refugia, rotations and sequences produce the same results. B),D) when costs are added there are infrequent model runs where rotations last longer below resistance thresholds than sequences. For between 6 and 7% of runs including costs, rotations last more than 20% longer than sequences. C) when dispersal from untreated refugia is included without costs there are even less frequent runs where rotations last longer than sequences (in just 4 out of 10000 runs rotations lasted greater than 20% longer than sequences)."}

@@

#now that non thershold runs are not removed :
#which(df_cost$pcent_rot_adv > 200)
#987 1023 1324 1365 1511 2051 2717 3187 3196 3582 3589 3768 4128 4219 4835 4880 6233 6538 6913 8666 9361
scenarios_to_plot <- c(2717)

@@

    gg <- ggplot( to_plot, aes_string(x='inputs',y='PRCC') ) +     
             geom_point(shape=1, colour='red') +
             theme(axis.text.x = axis_text_x) +
             geom_hline(yintercept = 0, linetype=3) +
             #theme(panel.grid.major.y = element_line(color = "grey", size = 1)) +
             ylim(-1,1) +
             #ggtitle(paste(strat_name[strategy_num])) +
             xlab(NULL)

@@

```{r Fig10, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9, fig.show='hold', fig.align='center', fig.cap="Figure 10. Simplified indication of how fitness differences between selective advantages and costs are expected to change according to dominance and resistance allele frequencies. Top left : high dominance of cost, low dominance of selection, difference between cost and selection fitness components can be very low at low allele frequencies, difference increases as allele frequencies increase. Lower left: low dominance of cost, low dominance of selection : increased difference at the lowest frequencies. Top right: increased dominance of selection markedly increases benefits of resistance at low frequencies and thus the difference between costs and benefits even when dominance of cost is high. This shows why dominance of selection needs to be low and dominance of cost needs to be high, when dominance of selection is high the difference between costs and benefits is high even at low frequencies irrespective of the value of dominance of cost. When dominance of selection is low a high dominance of cost can lead to very small differences between costs and benefits at low frequencies. In these latter circumstances if rotations can keep frequencies low they can reduce the long term development of resistance. Plot calculation assumes Hardy-Weinberg equilibrium which will not be the case under selection but the general pattern will be similar. Fitness cost set to 0.05 and selective advantage of resistance to 0.4."}

>>>>

*todo add to this description of how the model works and refer to SI where we can show the algebra*

Within the model, the inputs described in Table 1 determine the fitness in each generation as shown in Figure 1, differences in fitness lead to the changes in allele frequencies in the following generation.

@@

*Ian : at this stage it would be good to revisit the following model details.*    
*It is fairly easy for me to rerun all the analyses if we change our mind on these.*

#### Optional model details    
Two optional model details were included to aid giving a more meaningful comparison of strategies. Firstly allele frequencies in a run were not allowed to go below the starting frequencies randomly chosen for that run. This aims to represent a situation where in the absence of insecticide use allele frequencies are maintained in balance between mutation and selection. Without this option enabled, in runs including costs of resistance, resistance allele frequencies can decline below starting levels when an insecticide is not in use and this makes it more difficult to compare alternate strategies. *Ian, can you help out with a few sentences here ?*

Secoundly a constraint was added to stop rapid switching between insecticides every few generations which would be impractical in the field. Certain input combinations create this behaviour in the absence of a constraint. For example when two insecticides are close to the resistance frequency threshold of 0.5 it is possible for them to switch between each other every few generations as frequencies go above the threshold when in use and decline below it when not in use. Two different ways of stopping this rapid switching behaviour were trialled. In one, a minimum-insecticide-use constraint stopped switching to another insecticide within a set number of generations (5 in this case) of starting using another. This could represent a field situation where once walls have been sprayed or nets distributed they cannot be unsprayed or recalled. In a second, a minimum-repeat-interval constraint stopped returning to the use of an insecticide within set number of generations (again we used 5 generations) since it was last used.     
*I'm looking into the implications of these different constraints now.*

@@

It is necessary to choose a measure to quantify the performance of an insecticide-use strategy and enable comparison with an alternative. For this we summed the number of generations when an insecticide was deployed and the resistance allele frequency for that insecticide was below 0.5. This allowed comparison between strategies using a variable number of insecticides.

@@

When resistance costs are included (Fig 3), resistance allele frequencies decline in periods when the corresponding insecticide is not in use. Thus for sequences the decline can allow insectides to be returned to use later on because resistance has declined below threshold levels in contrast with the situation in Figure 2.

Adding representation of an untreated refugia and dispersal between it and the treated area also allows resistance allele frequencies to decline when an insecticide is not in use (Fig 4). Resistance frequencies in the treated areas (red lines) increase when an insecticide is in use. Resistance frequencies in the untreated areas (blue lines) increase after those in the treated areas as a result of dispersal from the treated areas. When an insecticide stops being used the resistance frequency in the treated area declines due to dispersal from the untreated area. For the sequence the resistance frequency for each insecticide increases to reach the threshold in one step and consistent with the previous figure resistance frequencies decline when the insecticide is not in use allowing it to be used again. For the rotation the resistance frequency for each insecticide increases in shorter steps while it is in use and similarly declines while not.

#### Results2 : comparing rotations and sequences across model runs.

@@

Thus to aid the comparison of rotations and sequences model runs can be broadly classified into four groups :

1. both strategies succeeded : for both rotations and sequences resistance thresholds for all insecticides not reached within 500 generations.
2. neither strategy succeeded
3. rotation only succeeded
4. sequence only succeeded

The number of model runs in each of these groups for the different scenarios are shown in Fig 5. In the base scenario with no costs or dispersal (Fig 5A) the majority of runs do not succeed, a small proportion succeed for both strategies and none succeed for either strategy alone. When costs of reistance are added (Fig 5B) more runs succeed for both strategies, up to about half of those that succeed for neither, and a small proportion (< 4%) succeed for rotations alone. Adding both costs and dispersal results in more runs succeeding for both than neither (Fig 5C) and the proportion of runs where rotations only succeed is similar to the scenario with costs alone.

The runs in each of these three main classes (success in both, neither or rotations only) differ in their distribution of input values (Fig 5xtra). Runs in which both strategies were successful were more likely to have higher costs and lower effectiveness, exposure and resistance restoration. Runs in which neither strategy was successful were the reverse. Runs for which rotations only were successful tended to have higher dominance of cost, lower dominance of resistance and lower rotation intervals.

@@

As shown in Fig 6B some runs with costs of resistance can show substantial advantages for rotations over sequences. Examples of one such run is shown in Fig 7.

A Partial Rank Correlation analysis identified that of the inputs, dominance of cost and dominance of selection had the greatest influence on the difference between rotations and sequences when resistance costs were included (Fig 8). A combination of high values of dominance of cost and low values for dominance of selection are required to give the greatest advantages to rotations, but do not guarantee this advantage (Fig 9).

@@

This can be understood intuitively. When there are no costs, the dynamics are the same, it is simply that sequential deployment proceeds in a few, larger steps while rotations proceeds in more numerous smaller steps; both reach the same point in around the same time. However dominance effects on cost and resistance can act to keep frequencies low in a rotations policy. Recall that at low frequencies of resistance most resistance alleles are present as heterozygotes RS and only a vey small proportions heterozygotes. However costs are dominant and still act at low frequencies. This is illustrated on Figure Y. Relatively rapid rotations keep the allele frequencies at the LH side of Figure 10 where costs during the periods when the insecticide is not deployed greatly outweigh the selection for IR during the periods when the insecticide is deployed. We believe the explanation is therefore that rotations keep allele frequencies in the parameter space to the left of Fig 10 and hence acts to slow the evolution of IR compared to sequences which allow frequencies to rise to relatively high levels before the insecticide is replaced. Note also that starting frequency, as expected, had an impact (Fig 8), lower starting frequencies slightly favouring rotations by allowing more time in the “low frequency” parameter space. [*Andy not sure that the effect of starting frequencies in the PRCC is notable, seems very close to the line of not having an effect either way*]

@@

1. *add to methods description of how fitness determined in model, move & modify fig 1*
1. *add caveat about polygenic resistance*
1. *talk about cross resistance*

@@

ggrot <- rot_plot_resistance(dfrot, plot_refuge=TRUE, logy=TRUE, plot=FALSE, title='A. Rotation')
ggseq <- rot_plot_resistance(dfseq, plot_refuge=TRUE, logy=TRUE, plot=FALSE, title='B. Sequence')
library(gridExtra)
grid.arrange(ggrot, ggseq, ncol=1)

@@

```{r Fig6, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=4, fig.show='hold', fig.align='center', fig.cap="Figure 6. Effect of cost and dispersal on the difference between rotations and sequences from model runs. The y axis shows the number of model runs for the different categories of difference between rotations and sequences. A) when there are no resistance costs or dispersal from untreated refugia, rotations and sequences produce the same results. B),D) when costs are added there are infrequent model runs where rotations last longer below resistance thresholds than sequences. For between 6 and 7% of runs including costs, rotations last more than 20% longer than sequences. C) when dispersal from untreated refugia is included without costs there are even less frequent runs where rotations last longer than sequences (in just 4 out of 10000 runs rotations lasted greater than 20% longer than sequences)."}

@@

#now that non threshold runs are not removed :
#which(df_cost$pcent_rot_adv > 200)
#987 1023 1324 1365 1511 2051 2717 3187 3196 3582 3589 3768 4128 4219 4835 4880 6233 6538 6913 8666 9361
scenarios_to_plot <- c(2717)

@@

    gg <- ggplot( to_plot, aes_string(x='inputs',y='PRCC') ) +     
             geom_point(shape=1, colour='red') +
             theme(axis.text.x = axis_text_x) +
             geom_hline(yintercept = 0, linetype=3) +
             theme_minimal() +
             #theme(panel.grid.major.y = element_line(color = "grey", size = 1)) +
             ylim(-1,1) +
             #ggtitle(paste(strat_name[strategy_num])) +
             xlab(NULL)

@@

```{r Fig10, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9, fig.show='hold', fig.align='center', fig.cap="Figure 10. Simplified indication of how fitness differences between selective advantages and costs are expected to change according to dominance and resistance allele frequencies. Top left : high dominance of cost, low dominance of selection, difference between cost and selection fitness components can be very low at low allele frequencies, difference increases as allele frequencies increase. Lower left: low dominance of cost, low dominance of selection : increased difference at the lowest frequencies. Top right: increased dominance of selection markedly increases benefits of resistance at low frequencies and thus the difference between costs and benefits even when dominance of cost is high. This shows why dominance of selection needs to be low and dominance of cost needs to be high, when dominance of selection is high the difference between costs and benefits is high even at low frequencies irrespective of the value of dominance of cost. When dominance of selection is low a high dominance of cost can lead to very small differences between costs and benefits at low frequencies. In these latter circumstances if rotations can keep frequencies low they can reduce the long term development of resistance. Note that with higher costs or lower benefits of resistance there is the potential for the cost and benefit lines to cross and for there to be a region at low frequencies where resistance is not selected for. Plot calculation assumes Hardy-Weinberg equilibrium proportion of genotypes. The equilibriun will not be followed in this case because the resistance locus is under selection but it serves as an approximation to illustrate the general effect. Fitness cost set to 0.05 and selective advantage of resistance to 0.4."}

### 2018-05-18 inst/documents/rotations_notes.Rmd
paper3

@@

Decide whether to use min_rwr_interval OR min_gens_switch_back

Before sending to Ian.

~ add to methods about min rwr interval
~ add about keeping resistance allele freq at starting level
~ add to fig Y legend Ians caveats

@@

@@

~ can I add figures from specific sensi_an runs ? or safer to run the model for set runs from the paper rmd.

~ how to sort library.bib ? Have a much reduced copy in rotations folder now.

@@

be careful that min_rwr_interval doesn't come back to bite me. Ian wanted it different, i.e. not going back to an insecticide that has been used recently ...

~ sensi_an_rotations1.Rmd

~ *decide on min_rwr_interval for sensi analysis
~ decide on max_gen, what was it for paper1 ?
~ one generation per month, 12 generations per year, 500 gens is ~40 years

~ maybe come back to i_in_use ribbons not meeting issue

~ rename rot_criterion to rot_threshold

~ add in some unit-tests ?

~ check how gens_under_thresh are counted in sensi_an_rotations1
    summarise(gens_dep_under50 = sum(resistance < 0.5 &
                                     #finds insecticide in use = this one
                                     resist_gene==paste0('insecticide',insecticide), na.rm=TRUE)) %>%
                                 summarise(tot_gens_dep_under50 = sum(gens_dep_under50)) %>%    
                                 unlist()

@@

Download papers on modelling oxitech
Insecticide resistance
https://doi.org/10.1093/jee/100.5.1642
https://doi.org/10.1603/029.102.0233

@@

Type of fitness cost influences the rate of evolution of resistance to transgenic Bt crops
Looks like could be very interesting :
DOI: 10.1111/1365-2664.12680

@@

~ check again on current small differences between rotations & sequences

~ maybe I can write a function to plot 2 scenarios on the same graph (in the same way that curtis fig2 plots mix & seq on same graph) ?
~ I could have the insecticide use still along the top but with diff colours showing the two alternatives

~ get sensi_an_rotations1.Rmd working for Ians Q1

@@

*Ian said first question is to assess whether there is any benefit to rotations when there are no costs and no emigration*

@@

~ create a new version of run_rot2() allowing me to refactor the code and check that it still generates same results as run_rot()

@@

# useful
~ one generation per month, 12 generations per year, 500 gens is ~40 years

>>>>

done ~ add to methods about min rwr interval
done ~ add about keeping resistance allele freq at starting level
done ~ add to fig Y legend Ians caveats

*14/5/2018*
~ Paper3, decsribing results and adding to methods clarifiying model constraints.

Before sending to Ian.
~ resize no square figs, make sure one per page

looking at how I made smaller reference bib file for uea work.

Seems I can export a single non-synced bibtex file

https://www.imperial.ac.uk/media/imperial-college/administration-and-support-services/library/public/LaTeX-and-Mendeley-sep-2017.pdf
1. Exporting a BibTeX (.bib) file from your Mendeley library
• Select the records for export
• From top menus, go to File>Export…
• In Popup box, choose a file name (no spaces); ensure file type is set as BibTeX
• Save the .bib file in the same folder as the LaTeX document

df_res_all_rot_nocost_nomig_10000_mgs5.rda

mgs5 to indicate min_gens_switch_back=5

Aha! all these big sequence advantages (well those >50%) 
were due to artfefact when rotation interval=5 and num insecticides=2 rotation couldn't return to first insecticide because it was only 5 generations ago !!

*17/5/2018*

New papers to cite in rotations paper :

Hackett (2016) 
https://www.ncbi.nlm.nih.gov/pmc/articles/PMC5026168/
interesting model of resistance to transgenic crops with similar results on the importance of dominance of resistance and cost :
"the most significant constraint upon model behaviour is the dominance of the resistant phenotype, with greater dominance decreasing Bt usage. For a given level of dominance of resistance, the type and dominance of any fitness costs may relax these constraints, allowing for larger areas of Bt crop [smaller refuges to control resistance evolution]"

The R code is here and looks quite accessible :
https://datadryad.org//resource/doi:10.5061/dryad.4s405

Allen(2017) Reversing resistance: different routes and common themes across pathogens
http://rspb.royalsocietypublishing.org/content/284/1863/20171619

Mostly applied to microbes & fungi, some good general points, probably not useful to us.
Does talk about different types of reversion and the importance of cost.

My sentence from an example about negative cross resistance :
Negative cross resistance meant that two fungicides were used successfully on Botrytis cinerea in European vineyards for 7 years ending in 1994 when a new mutation appeared that gave resistance to both. [87]

Russell PE. 1995 Fungicide resistance—occurrence and management. J. Agric. Sci. 124, 317–323. (doi:10.1017/S0021859600073275)

So negative cross resistance can work for a while but is susceptible to new mutations giving resistance to both ingredients.

Alphey & Bonsal 2017 good accessible description of high dose-refuge strategy :
Genetics‐based methods for agricultural insect pest management
(but the paper itself is about sterile insect technique)

For example, modelling the effects of larval competition and exploring late‐acting lethal phenotypes in mosquitoes predicted that this could be substantially more effective at population control than an early‐acting (e.g. embryonic) lethality or radiation‐induced sterility (Atkinson et al., 2007; Phuc et al., 2007; Alphey & Bonsall, 2014a). Indeed, if density‐dependent juvenile competition were over‐compensatory, genetic lethality that occurred at an earlier stage, thereby freeing survivors from regulation by intense competition, could push adult insect numbers higher than in the natural uncontrolled population.

In terms of the genetics, consider a resistant allele r, which is initially rare. The dominant allele S is susceptible to Bt. If the high dose assumption is achieved, only some rr individuals can survive a full life cycle on transgenic plants and emerge as adults to mate. In the refuge, most emerging adults will be susceptible SS, especially if the r allele has fitness costs in the absence of the Bt toxin. If the refuge is located so that the two subpopulations are well‐mixed, most resistant rr survivors will mate with susceptible SS insects from the refuge. Their resulting Sr progeny cannot survive on the Bt crops and so will not pass on the resistant allele to future generations. Unfortunately, a few insect species, such as the economically important pests Helicoverpa zea (Boddie) and Helicoverpa armigera (Hübner) (both species, confusingly, known as both bollworm and corn earworm), have been identified in the past as ‘moderate dose pests’, where the toxins were unable to kill all heterozygotes (EPA, 1998; Tabashnik et al., 2008). Even where its main assumptions appear to hold, the high‐dose/refuge strategy is predicted only to delay resistance and, after two decades of commercially grown Bt cotton and Bt corn (maize), some field‐evolved resistance has now been observed and reported in a variety of insect species. In some cases, this has already led to reduced efficacy of crops, or even crop failures. A comprehensive review is provided by Tabashnik et al. (2013).

In economic terms, the high‐dose/refuge strategy is an inter‐temporal trade‐off, sacrificing current value to retain value generated in future (Frisvold & Reeves, 2008). Refuge plants will be damaged, which reduces their yield if they are crops, or reduces the area allotted to crop production if alternative host plants are used. This damage or lost production is tolerated, in return for prolonging the efficacy of the protection afforded by the Bt crops. In principle, a conventional crop refuge could be sprayed with another pesticide (one with no cross‐resistance between its active ingredient and Bt), although doing so reduces its effectiveness as a refuge.

Aha and a justification for using resistance threshold of 0.5 !
Generally, a frequency of 0.5 is an indicator of a serious problem because resistance frequency increases rapidly once r becomes the more common allele (Carrière & Tabashnik, 2001).

Carrière & Tabashnik, 2001. Modelling refuges in Bt corn
Good paper to cite in rotations paper.
They found some similar things to us, e.g. resistance not increasing when dominance of resistance low and dominance of cost high.

*Carriere 2018* good one to include, recent data on fitness costs showing that there are highish fitness costs but that they are almost completely recessive.
Genotype-specific fitness cost of resistance to Bt

*18/5/2018*

*Glunt 2018* Impact of temperature on insecticide toxicity.
Relevant to window of selection. 
In arabiensis & funestus with deltamethrin, generally mortalities higher when temperatures either higher or lower than standard. The difference between resistant and susceptible mortality is generally less, suggesting that the window of selction may be lower outside of standard lab conditions. Bendiocarb mortality increased with temperature and unsurprisingly was similar between strains susceptible and resistant to pyrethroids.

@@

*today*
done ~ reread Ians text about dominance
~ move rotations repo back to my github
~ update R
done ~ tweet to malaria journal
~ think more about rotations datavis
done ~ read the alphey oxitech modelling papers

To talk with Ian about next week :
~ min_rwr_interval vs min_gens_switch_back

Look at Hackett R code. Could it be an even simpler model that we could use ?
The R code is here and looks quite accessible :
https://datadryad.org//resource/doi:10.5061/dryad.4s405

Looking at moving rotations repo back to me. Seems transfer may be best way :
https://help.github.com/articles/about-repository-transfers/
Transfer is at the bottom of settings, options : from Ians account

Look into whether to use min_rwr_interval OR min_gens_switch_back 
I'm thinking that I'll stick with min_rwr_interval

Revisit Ians paragraph about the rare conditions that give large benefit to rotations.

"Note however that this combination of recessive resistance and high dominance costs are exactly the features most desirable in any insecticide irrespective of how it is deployed. ... it is very difficult to maintain insecticide concentration at the level required to make resistance recessive. Concentrations decay post-application in most deployment strategies and this cause resistance to gradually change from being recessive to being dominant.""
me is there anything similar on the dominance of costs ?

@@

In rotations and WoS MS we should talk about high dose/refuge strategy for transgenic crops (specifically Bt corn). The high dose part is that the dose is sufficiently high to kill all heterozygotes i.e. making resistance recessive. But if this was used across a large area the evolution of resistance would be very rapid - because only homozygous resistant individuals would be left. The refuge part is to provide a source of susceptible immigrants which can generate heterozygous offspring that are vulnerable to the dose.

@@

@@

~ *decide on min_rwr_interval for sensi analysis
~ one generation per month, 12 generations per year, 500 gens is ~40 years

~ maybe come back to i_in_use ribbons not meeting issue
~ rename rot_criterion to rot_threshold
~ add in some unit-tests ?
~ check how gens_under_thresh are counted in sensi_an_rotations1
    summarise(gens_dep_under50 = sum(resistance < 0.5 &
                                     #finds insecticide in use = this one
                                     resist_gene==paste0('insecticide',insecticide), na.rm=TRUE)) %>%
                                 summarise(tot_gens_dep_under50 = sum(gens_dep_under50)) %>%    
                                 unlist()

@@

@@

@@

@@

@@

@@

# useful
~ one generation per month, 12 (or 10) generations per year, 500 gens is ~40-50 years
